IJAR Bayesian Nonparametrics Special Issue Manuscript

### 2015-10-20 ijar-npb-ms.tex
extended set of prior workings in sec4, minor changes in rest

$p \mid s$ is Beta distributed with the updated parameters
\begin{align*}
\nn &= \nz + n\,, &
\yn &= \frac{\nz}{\nz + n} \cdot \yz + \frac{n}{\nz + n} \cdot \frac{s}{n}\,,
\end{align*}
so after seeing that $s$ out of $n$ components function (at time $t$),
the posterior mean $\yn$ for $p$ is a weighted average of
the prior mean $\yz$ and $s/n$ (the fraction of function components in the data),
with the weights $\nz$ and $n$, respectively.
We see that $\nz$ plays the same role for the prior mean $\yz$
as the sample size $n$ for the observed mean $s/n$,
thus the notion of pseudocount.
Also, the higher $\nz$, the higher the weight for $\yz$
in the weighted average calculation of $\yn$,
so $\nz$ gives the strength of the prior as compared to the sample size $n$.
%The parameters $\nz$ and $\yz$ have a more intuitive interprtation,

@@

As shown in Walter \& Augustin (2009),
we can have both tractability and meaningful reaction to prior-data conflict
by using sets of priors $\MktZ$ produced by parameter set $\PktZ = [\nktzl, \nktzu] \times [\yktzl, \yktzu]$. 
For such a rectangular prior parameter set $\PktZ$,
the posterior parameter set $\PktN$ is not rectangular anymore.

In case of no prior-data conflict ($s^k_t/n^k_t \in [\yktzl, \yktzu]$),
the set shrinks in the $\ykt$ dimension
(how much it shrinks depends on $\nktz$, leading to the so-called spotlight shape),
leading to smaller $\Rsys(t)$ intervals.

In case of prior-data conflict ($s^k_t/n^k_t \not\in [\yktzl, \yktzu]$),
$\PktN$ has the so-called banana shape,
giving a wider $\yktn$ interval as compared to the non-conflict case. 
As a consequence, we will have wider intervals for $\Rsys(t)$ in case of prior-data conflict.

Lower and upper bounds for $\Rsys(t)$ by min and max over $\PZ_{1,t}, \ldots, \PZ_{K,t}$.

Monotonicity in $\yz_{k,t}$, so we get
\begin{align*}
\lRsys(t)
 &= \min_{\nz_{1,t},\ldots,\nz_{K,t}} \Rsys(t) \\
 &= \min_{\nz_{1,t},\ldots,\nz_{K,t}} 
    \sum_{l_1=0}^{m_1} \cdots \sum_{l_K=0}^{m_K} \Phi(l_1, \ldots, l_K)
                                                 \prod_{k=1}^K P(C^k_t = l_k \mid s^k_t) \\
 &= \min_{\nz_{1,t},\ldots,\nz_{K,t}} 
    \sum_{l_1=0}^{m_1} \cdots \sum_{l_K=0}^{m_K} \Phi(l_1, \ldots, l_K) \times \\ & \hspace*{12ex}
    \prod_{k=1}^K {m_k \choose l_k} \frac{B(l_k + \nn_{k,t}\ynl_{k,t}, m_k - l_k + \nn_{k,t}(1-\ynl_{k,t}))}
                                         {B(\nn_{k,t}\ynl_{k,t}, \nn_{k,t}(1-\ynl_{k,t}))} \,.
\end{align*}

@@

\section{Computations}

***publish code as seperate \textbf{R} package, or include in Louis' package \texttt{ReliabilityTheory}
(which contains the function to calculate the survival signature)***

\section{Example}

***Will ask Simme Douwe Flapper (Prof at TU Eindhoven) for real-world example,
otherwise some toy example will do.***

Show $\Rsys(t)$ intervals for prior-data conflict and non-conflict data
to show effect of prior-data conflict.

Also calculate $\Rsys(t)$ intervals for vacuous prior ($[\yktzl, \yktzu] = [0,1]$, some fixed $\nktz$)?***

Show the effect of extra redundancy on $\Rsys(t)$ intervals like in Risk Analysis paper,
or show effect of replacement of failed components in a system
(replaced components constitute a new type with shifted component reliability function)

@@

Extend the model to deal with right-censored observations which are common in realibility setting.
We assume that a minimal assumption (component can fail immediately after censoring or live forever)
will be simple to implememt but will lead to high imprecision,
whereas assuming exchangeability with other components at moment of censoring
will be more complex to accomodate but will lead to les imprecision.

>>>>

$p \mid s$ is Beta distributed with the updated parameters
\begin{align}
\nn &= \nz + n\,, &
\yn &= \frac{\nz}{\nz + n} \cdot \yz + \frac{n}{\nz + n} \cdot \frac{s}{n}\,,
\label{eq:nyupdate}
\end{align}
so after seeing that $s$ out of $n$ components function (at time $t$),
the posterior mean $\yn$ for $p$ is a weighted average of
the prior mean $\yz$ and $s/n$ (the fraction of function components in the data),
with the weights $\nz$ and $n$, respectively.
We see that $\nz$ plays the same role for the prior mean $\yz$
as the sample size $n$ for the observed mean $s/n$,
thus the notion of pseudocount.
Also, the higher $\nz$, the higher the weight for $\yz$
in the weighted average calculation of $\yn$,
so $\nz$ gives the strength of the prior as compared to the sample size $n$.
%The parameters $\nz$ and $\yz$ have a more intuitive interprtation,

@@

As was shown in Walter \& Augustin (2009),
we can have both tractability and meaningful reaction to prior-data conflict
by using sets of priors $\MktZ$ produced by parameter sets $\PktZ = [\nktzl, \nktzu] \times [\yktzl, \yktzu]$. 
In our model, each prior parameter pair $(\nktz, \yktz) \in \PktZ$
corresponds to a Beta prior, thus $\MktZ$ is a set of Beta priors.
The set of posteriors $\MktN$ is obtained by updating each prior in $\MktZ$ according to Bayes' Rule.
Due to conjugacy, $\MktN$ is a set of Beta distributions with parameters $(\nktn, \yktn)$
obtained from updating $(\nktz, \yktz) \in \PktZ$ according to \eqref{eq:nyupdate},
and so can be described efficiently through the set of updated parameters
\begin{align*}
\PktN &= \Big\{ (\nktn, \yktn) \mid (\nktz, \yktz) \in \PktZ = [\nktzl, \nktzu] \times [\yktzl, \yktzu] \Big\}\,.
\end{align*}
Such rectangular prior parameter sets $\PktZ$ have been shown
to balance desirable model properties and ease of elicitation quite well.

For each component type $k$ and time point $t$,
one has to choose only the four parameters $\nktzl, \nktzu, \yktzl, \yktzu$
(so $4 \times |{\cal T}|$ parameters are needed to define the set of prior distributions characterizing
prior knowledge on the survival function of components of type $k$);
notwithstanding this, the model has favourable inference properties.

Central to these inference properties is that
while the prior parameter set $\PktZ$ is rectangular,
the posterior parameter set $\PktN$ is not rectangular anymore;
indeed, the shape of $\PktN$ depends on the presence of prior-data conflict,
which is naturally operationalized as $s^k_t/n^k_t \not\in [\yktzl, \yktzu]$:
prior-data conflict is present when the observed fraction of functioning components
is outside its expected range.

In absence of prior-data conflict, 
$\PktN$ shrinks in the $\ykt$ dimension;
how much it shrinks depends on $\nktz \in [\nktzl, \nktzu]$,
leading to the so-called spotlight shape depicted in Figure***.
As $\yktn$ gives the posterior expectation for the functioning probability $p_t^k$,
shorter $\yktn$ intervals mean more precise knowledge about $p_t^k$.
Also, the variance interval for $p_t^k$ will decrease
as the Beta distributions in $\MktN$ will be more pointed
due to the increase of $\nktz$ to $\nktn$.s

In case of prior-data conflict ($s^k_t/n^k_t \not\in [\yktzl, \yktzu]$),
$\PktN$ has instead the so-called banana shape,
arising from $\yktn$ intervals being shifted closer towards $s^k_t/n^k_t$
for lower $\nktn$ values than for higher $\nktn$ values, see Figure***.
Overall, this results in a wider $\yktn$ interval as compared to the non-conflict case, 
reflecting the extra uncertainty due to prior-data conflict,
thus making more cautious probability statements about $p_t^k$.

***Lower and upper bounds for $\Rsys(t)$ by min and max over $\PtZi{1}, \ldots, \PtZi{K}$.

***Monotonicity in $\yz_{k,t}$, so we get, for each $t \in {\cal T}$,
\begin{align*}
\lefteqn{\lRsys(t \mid \vec{t}^1, \ldots, \vec{t}^K)}\\
 &= \min_{\PtZi{1}, \ldots, \PtZi{K}} \Rsys(t \mid \PtZi{1}, \ldots, \PtZi{K}, \vec{t}^1, \ldots, \vec{t}^K) \\
 &= \min_{\nz_{1,t},\ldots,\nz_{K,t}} 
    \sum_{l_1=0}^{m_1} \cdots \sum_{l_K=0}^{m_K} \Phi(l_1, \ldots, l_K)
                                                 \prod_{k=1}^K P(C^k_t = l_k \mid \yktzl, \nktz, s^k_t) \\
 &= \min_{\nz_{1,t},\ldots,\nz_{K,t}} 
    \sum_{l_1=0}^{m_1} \cdots \sum_{l_K=0}^{m_K} \Phi(l_1, \ldots, l_K) \times \\ & \hspace*{12ex}
    \prod_{k=1}^K {m_k \choose l_k} \frac{B(l_k + \nn_{k,t}\ynl_{k,t}, m_k - l_k + \nn_{k,t}(1-\ynl_{k,t}))}
                                         {B(\nn_{k,t}\ynl_{k,t}, \nn_{k,t}(1-\ynl_{k,t}))} \,.
\end{align*}

***How to choose the sets of priors for the survival function of type $k$ components,
i.e. how to choose $\PkZi{1}, \ldots \PkZi{\tmax}$?***

@@

\section{Computations \& Toy Example}

***publish code in Louis' package \texttt{ReliabilityTheory}
(which contains the function to calculate the survival signature)***

***Show $\Rsys(t)$ intervals for prior-data conflict and non-conflict data
to show effect of prior-data conflict.***

***Also calculate $\Rsys(t)$ intervals for vacuous prior ($[\yktzl, \yktzu] = [0,1]$ for some fixed $\nktz$)?***

***Show the effect of extra redundancy on $\Rsys(t)$ intervals like in Risk Analysis paper,
or show effect of replacement of failed components in a system
(replaced components constitute a new type with shifted component reliability function)?***

@@

***Extend the model to deal with right-censored observations which are common in the relaibility setting.
We assume that a minimal assumption (component can fail immediately after censoring or live forever)
will be simple to implememt but will lead to high imprecision,
whereas assuming exchangeability with other surviving components at moment of censoring
will be more complex to accomodate but will lead to less imprecision.
